---
title: "data_cleaning"
author: "Yuxuan Chen"
date: "12/1/2021"
output: github_document
---

Import the libraries
```{r, message=FALSE}
library(tidyverse)
library(rvest)
library(httr)
library(maps)
library(readxl)
library(plotly)
library(choroplethr)
library(choroplethrMaps)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# data cleaning

**Part I: By state**

We've found that instead of having the usual 50 states, Washington DC, and New York City, the data set also contains data for the whole US. We choose to focus on the 52 jurisdictions (including the 50 states, DC, and NYC) at first.

```{r message = FALSE}
drug_overdose = read_csv("./data/VSRR_Provisional_Drug_Overdose_Death_Counts.csv") %>% 
  janitor::clean_names()

unique(pull(drug_overdose, state))
# contains 50 states, Washington DC, whole US, and the City of New York.
```

 - State level: 
```{r}
state_level = c(state.name[1:8], "District of Columbia", state.name[9:32],"New York City", state.name[33:50])
```


```{r}
drug_overdose_52 = 
  drug_overdose %>% 
  filter(!(state_name %in% c("United States"))) %>% 
  relocate(state_name) %>% 
  mutate(month = factor(month, levels = month.name), # change month and year to factor
         year = factor(year),
         state_name = factor(state_name, levels = state_level)) %>% 
  arrange(state_name) %>% 
  group_by(state_name, year) %>% 
  mutate(month = sort(month)) # sort by month order
```

 - Since the dataset mostly doesn't have the death count values for specific types of drug overdose, for this dataset, we only choose the data_value with indicator of `Number of Deaths`, `Percent with drugs specified`, and `Number of Drug Overdose Deaths`.

```{r}
# (For Plot): 
# percentage of drug overdose deaths over total number of deaths
# percentage of drugs specified (text describe)

drug_overdose_death = 
  drug_overdose_52 %>% 
  select(-c(state, footnote_symbol, percent_complete, period, percent_pending_investigation, predicted_value)) %>% 
  filter(indicator %in% c("Number of Deaths", "Percent with drugs specified", "Number of Drug Overdose Deaths"))
drug_overdose_death
```

 - `Number of Drug Overdose Deaths` is predicted since the `Percent with drugs specified` is not 100%. 
 

footnote categorization: I AM STILL NOT SURE OF THIS PART. I JUST KEPT CREATING DUMMY VARIABLES!!
```{r}
#unique(pull(drug_overdose_51, footnote))
drug_overdose_52 = 
  drug_overdose_52 %>%
  mutate(low_data_quality = ifelse(str_detect(footnote, "low data quality"), 1, 0), # data_value not shown, predicted yes?
         suppressed = ifelse(str_detect(footnote, "suppressed"), 1, 0),
         underreported = ifelse(str_detect(footnote, "Underreported"), 1, 0)) %>% 
  relocate(footnote, .after = last_col())
```

 
**drug categories: letter-number indicators**

Drug overdose deaths are identified using underlying cause-of-death codes from the Tenth Revision of ICD (ICD–10): X40–X44 (unintentional), X60–X64 (suicide), X85 (homicide), and Y10–Y14 (undetermined). Drug overdose deaths involving selected drug categories are identified by specific multiple cause-of-death codes.

Drug categories presented include: heroin (T40.1); natural opioid analgesics, including morphine and codeine, and semisynthetic opioids, including drugs such as oxycodone, hydrocodone, hydromorphone, and oxymorphone (T40.2); methadone, a synthetic opioid (T40.3); synthetic opioid analgesics other than methadone, including drugs such as fentanyl and tramadol (T40.4); cocaine (T40.5); and psychostimulants with abuse potential, which includes methamphetamine (T43.6). Opioid overdose deaths are identified by the presence of any of the following MCOD codes: opium (T40.0); heroin (T40.1); natural opioid analgesics (T40.2); methadone (T40.3); synthetic opioid analgesics other than methadone (T40.4); or other and unspecified narcotics (T40.6). This latter category includes drug overdose deaths where ‘opioid’ is reported without more specific information to assign a more specific ICD–10 code (T40.0–T40.4) (3,4). Among deaths with an underlying cause of drug overdose, the percentage with at least one drug or drug class specified is defined as that with at least one ICD–10 multiple cause-of-death code in the range T36–T50.8.

```{r}
drug_categories = 
  drug_overdose_52 %>%
  ungroup() %>% 
  select(-c(state, footnote_symbol, percent_complete, period, percent_pending_investigation, footnote, predicted_value)) %>% 
  filter(str_detect(indicator, "T4"))

# missing states' data:
drug_overdose_52 %>% 
  ungroup() %>% 
  select(state_name) %>% 
  unique() %>% 
  filter(!(state_name %in% drug_categories$state_name))
```

 - missing data in 9 states: Alabama, Arkansas, Florida, Idaho, Louisiana, Minnesota, Nebraska, North Dakota, Pennsylvania.

**Part II: counties in New York**

**NEED TO FIND DATA FOR 2019-2021**
```{r}
# for consistency between datasets, I only chose the data from 2015 to 2021

# model-based death rate:
predicted_deathrate_bycounty = 
  read_csv("./data/NCHS_-_Drug_Poisoning_Mortality_by_County__United_States.csv") %>% 
  janitor::clean_names() %>% 
  arrange(state) %>% 
  filter(year %in% (2015:2018),
         state %in% "New York")

predicted_deathrate_bycounty
# Plot by county 
# or combine, plot by state
```

```{r}
##unique(pull(predicted_deathrate_bycounty, census_division))
```

**Part III: the place that drug overdose happens (death place; injury place)** "drug_deaths"

```{r}

```

**Part IV: OPEN FOR DISCUSSION**

The following is "Completeness of 12-month ending provisional counts of **drug overdose deaths** relative to final counts from 2019 after six month lag, by reporting jurisdiction and ending month" from the CDC website. 
(check same datasets)

```{r}
url = "./data/ok.html"
drug_use_html = 
  read_html(url)  # r has imported the html page

 # we got the completeness data
completeness_52 = 
  drug_use_html %>% 
  html_table()  %>% 
  .[[2]] %>% 
  janitor::clean_names() %>% 
  rename(state_name = reporting_jurisdiction) %>% 
  filter(!(state_name %in% c( "United States"))) %>% 
  pivot_longer(jan:dec, names_to = "month", values_to = "completeness") %>% 
  mutate(month = month.name[match(str_to_title(month), month.abb)],
         month = factor(month, levels = month.name))
completeness_52
```
```{r}
# try to write a function to apply the completeness value to number_of_death_52.
data_imputation = function(original, criteria){
  original = original * ((100 - criteria)/100 + 1)
}
a = drug_overdose_death %>% 
  filter(str_detect(indicator, "Drug Overdose Deaths")) %>% 
  mutate(state_name = factor(state_name, levels = state_level)) %>% 
  filter(year == 2019) %>% 
  select(state_name, year, month, data_value) %>% 
  arrange(state_name, month) %>% 
  ungroup() %>% 
  mutate(new_value = map2_dbl(.x = data_value, .y = pull(completeness_52, completeness), ~data_imputation(.x, .y)))

a %>% 
 # plot_ly(x = ~neighbourhood, y = ~n, color = ~neighbourhood, type = "bar", colors = "viridis")
  plot_ly(x = ~month, y = ~new_value, color = ~state_name, type = "scatter",mode = "markers") %>% 
  add_lines(showlegend=FALSE)
```


**Part V: one representative state per census regions**

Northeast: NYC; Midwest: Ohio; South: Florida; West: California<br>

It's hard to find drug-specific data for those state missing specific drug types. Opioids data is much easier to find. We should probably focus on opioids.<br> 
We drew a map of the opioids to justify our choice.
```{r}
opioids_df = 
  drug_overdose_52 %>%  
  ungroup() %>% 
  select(state_name, year, month, indicator, data_value) %>% 
  filter(!(state_name %in% c("Alabama", "Arkansas", "Florida", "Idaho", "Louisiana", "Minnesota", "Nebraska", "North Dakota", "Pennsylvania")),
         str_detect(indicator, "T4")) %>% 
  mutate(opioids_yn = ifelse(str_detect(indicator, "opioids"), TRUE, FALSE)) %>% 
  group_by(year, month, opioids_yn) %>% 
  summarize(opioids_rate = mean(data_value, na.rm = TRUE))
opioids_df %>% 
  ungroup() %>% 
  ggplot(aes(x = month, y = opioids_rate, group = opioids_yn, color = opioids_yn))+
  geom_point()+
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1))+
  facet_grid(.~year)
```
<br>If we decided to focus on opioids, this will make my life much easier.

FL part:
Overall review:
```{r}
fl_death = 
  drug_overdose_52 %>% 
  filter(state_name %in% "Florida",
         indicator %in%  c("Number of Deaths", "Number of Drug Overdose Deaths")) %>% 
  select(year, month, indicator, deaths = data_value) %>% 
  pivot_wider(
    names_from = indicator,
    values_from = deaths
  ) %>% 
  janitor::clean_names() %>% 
  group_by(year, month) %>% 
  mutate(
    percent_overdose_death = number_of_drug_overdose_deaths / number_of_deaths
  )
 
```
```{r}
fl_death  %>% 
  ggplot(aes(x = month, y = percent_overdose_death, group = year, color = year))+
  geom_point()+
  geom_line()
```
```{r}
fl_death %>% 
  ggplot(aes(x = month, y = percent_overdose_death, group = NA))+
  geom_point()+
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1))+
  facet_grid(.~year)
```

We can also focus just on opioids 
Data source: cdc wonder, request by queries, Drug overdose deaths are identified using underlying cause-of-death codes from the Tenth Revision of ICD (ICD–10): X40–X44 (unintentional), X60–X64 (suicide), X85 (homicide), and Y10–Y14 (undetermined). Opioids identified by selecting T40.0-40.4, 40.6
```{r, message=FALSE, warning=FALSE}
# read in data
fl_opi_death = read_csv("./data/fl_opi_death_15_19.csv") %>% 
  janitor::clean_names() %>% 
  select(state_name = state, year, month, deaths)

fl_opi_death_add = read_csv("./data/fl_opi_death_20_21.csv") %>% 
  janitor::clean_names() %>% 
  select(state_name = occurrence_state, year = year_code, month, deaths) 

fl_opi_death = bind_rows(fl_opi_death, fl_opi_death_add) %>% 
  filter(!is.na(month)) %>% 
  separate(month, into = c("month", "useless"), sep = "\\,") %>% 
  select(-useless) %>% 
  mutate(month = substr(month, 1,3),
  month = month.name[match(str_to_title(month), month.abb)],
  month = factor(month, levels = month.name),
  year = factor(year)) %>% 
  select(-state_name)
```
data could be suppressed as indicated by the cdc website, the data value also supported by florida government

But anyway let's plot<br>
First by year, this is the death count
```{r}
fl_opi_death  %>% 
  ggplot(aes(x = month, y = deaths, group = year, color = year))+
  geom_point()+
  geom_line()
```


what about each month within a year
```{r}
fl_opi_death %>% 
  ggplot(aes(x = month, y = deaths, group = NA))+
  geom_point()+
  geom_line()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust= 1))+
  facet_grid(.~year)
```
It's climbing, every single month. The 2020 data is less probably because it's provisional and some data still under investigation

Let's analyze the number of death vs median household income
```{r}
# combine them!
fl_eco_df = 
  read_csv("./data/median_household_income_fl.csv") %>% 
  janitor::clean_names() %>% 
  select(year, household_income_by_race, household_income_by_race_moe, geography) %>% 
  filter(
    str_detect(geography, "FL|United States|Florida"),
    year >= "2015") %>% 
  mutate(year = factor(year))

fl_us_df = fl_eco_df %>% 
  filter(geography %in% c("United States", "Florida"))

fl_instate_df = fl_eco_df %>% 
  filter(str_detect(geography, "FL|Florida"))

fl_us_df %>% 
  mutate(text_label = str_c("Year: ", year, "\nMedian Household Income: $", household_income_by_race, 
                            "\nMargin of error: ± $", household_income_by_race_moe)) %>% 
  plot_ly(
    x = ~year, y = ~household_income_by_race, color = ~geography, text = ~text_label, 
    alpha = 0.5, type = "scatter", mode = "markers+lines", colors = "viridis", error_y = ~list(array = household_income_by_race_moe)) %>% 
  layout(
    title = "Median Household Income: FLorida vs. The U.S",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Median Household Income"))

fl_instate_df %>% 
  mutate(text_label = str_c("Year: ", year, "\nMedian Household Income: $", household_income_by_race, 
                            "\nMargin of error: ± $", household_income_by_race_moe)) %>% 
  plot_ly(
    x = ~year, y = ~household_income_by_race, color = ~geography, text = ~text_label, 
    alpha = 0.5, type = "scatter", mode = "markers+lines", colors = "viridis", error_y = ~list(array = household_income_by_race_moe)) %>% 
  layout(
    title = "Median Household Income: In State Comparison",
         xaxis = list(title = "Year"),
         yaxis = list(title = "Median Household Income"))
```
Income and drug overdose death percent , by year
```{r}
#patchwork
income_drug_df = 
  fl_death %>% 
  ungroup() %>% 
  group_by(year) %>% 
  summarize(overdose_death_rate = sum(number_of_drug_overdose_deaths)/sum(number_of_deaths)) %>% 
  inner_join(., fl_us_df %>% filter(geography %in% "Florida"))
income_drug_df %>% 
  ggplot(aes(x = household_income_by_race, y = overdose_death_rate, group = NA, color = year))+
  geom_point()+
  geom_line()
```
maps maps maps
```{r}
fl_county_df = 
  read_csv("./data/NCHS_-_Drug_Poisoning_Mortality_by_County__United_States.csv") %>% 
  janitor::clean_names() %>% 
  filter(state %in% "Florida",
         year >= 2015) %>% 
  select(year, county, population, death_rate = model_based_death_rate) %>% 
  separate(county, into = c("county", "useless"), sep = " County") %>% 
  select(-useless) %>% 
  mutate(year = factor(year),
         county = str_to_lower(county)) %>% 
  relocate(county)
fl_county_df
```
```{r}
#data(df_pop_county)
data(county.fips)
abc = county.fips %>% 
  separate(polyname, into = c("state", "county"), sep = "\\,") %>% 
  filter(state %in% "florida") %>% 
  select(-state) %>% 
  as.tibble()
fl_county_df = left_join(fl_county_df,abc, by = "county") %>% 
  select(county, year, death_rate, fips) %>% 
  filter(year == 2015)
fips_add = c(12027, 12091, 12109, 12111)
a = 1
for (i in c(13,46,55,56)){
  fl_county_df[i,4] = fips_add[a]
  a = a+1
}
fl_county_df %>% 
  group_by(fips) %>% 
  mutate(fips = as.numeric(fips)) %>% 
  rename(region = fips,
         value = death_rate) %>% 
  county_choropleth(state_zoom = c("florida"))
```


What about Ohio?
Well, it' similar to NYC and CA, i can use my spare time to copy clarie's code for NYC
```{r}
ohio_df = 
  drug_overdose_52 %>% 
  filter(state_name %in% "Ohio") 
```
```{r}
age_fhp %>% 
  group_by(age_group, large_fhp) %>% 
  ggplot(aes(x = fhp_size_mm, y = age_group, fill = age_group))+
  geom_density_ridges(scale = 0.8) +
  stat_density_ridges(quantile_lines = TRUE, quantiles = 2)+
  geom_vline(xintercept= 40, linetype='longdash', col = "red")+
  geom_text(mapping = aes(x = 40,y = 0,label = "FHP = 40 mm", hjust = -.1, vjust = -5), color = "red") +
  ggtitle("Distribution of FHP size Across Age Groups") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(y = "Age Group (years)", x = "FHP size (mm)")
```


